### 2016-06-29 text/old/appendix_DSLTrans.tex
Move old sections. Divorce paper from definitions file.

@@

Let us now turn our attention to the apply function in \cref{def:apply_function}. Its role is extend all model fragments found by matching a rule on a given input-output model, such that each of those fragments becomes isomorphic to the complete rule (minus its backward links). This process effectively creates the new objects and relations specified in the apply part of the rule, for each of the fragments found when matching the rule.

@@

@@

While the execution of the rules belonging to a layer happens in parallel, the execution of the layers of a transformation happens sequentially. As per \cref{def:transformation_step_semantics}, the input-output model $m_{inter}$ is the output of executing a given layer that is passed onto the next layer as input. Note that an empty input-output model ($\epsilon$) is passed as the second argument to the $layerstep$ relation in \cref{def:transformation_step_semantics}. This is because in \cref{def:layer_step_semantics} of layer step semantics, the second argument of the relation is used as an accumulator for the model fragments that are added to be added to the output of the previous layer once all the rules in the current layer have been executed. The transformation rules in a layer are expanded before execution ($l^{\star}$) such that polymorphism in the match elements can be handled (see \cref{def:transformation_rule_expansion_appendix}).

>>>>

\reviewer{Def. B.12. First of all, I wonder why you define match as a function rather than
just defining what a match is; namely, an element of the set you are constructing
here. Secondly (and more importantly), why do you not use the notion of homo-
morphism instead of explicitly talking about ismorphic subgraphs? Your way,
you lose the connection between glue and $||rl||$: you know they are isomorphic,
but what if $||rl||$ has symmetries - then there is more than one match, which in
your definition cannot be distinguished.}

\reviewer{You pay the price for this loss of information in Def. B.13, which now is broken: $a\Delta$ depends on $m_{glue} $(which was called glue in Def. 12, why the change in
notation?) in a way depending on the mapping from $m_{glue}$ to $rl$ which you have
discarded.}

\reviewer{Def. B.13 is broken in another way as well: it does not properly define E' but
merely imposes some requirements that are also fulfilled by choosing E' = E.
Instead I know from the discussion about traceability links in Section 4 (which
logically comes after this appendix!) that you really want to add such links for
all newly created nodes of the output graph to all pre-existing nodes of the input
graph; but that is not made clear anywhere at this point.}

@@

Let us now turn our attention to the apply function in \cref{def:apply_function}. Its role is to extend all model fragments found by matching a rule on a given input-output model, such that each of those fragments becomes isomorphic to the complete rule (minus its backward links). This process effectively creates the new objects and relations specified in the apply part of the rule, for each of the fragments found when matching the rule.

@@

\reviewer{Def. B.14: It seems to me rather roundabout to sequentially apply all rules, ac-
cumulate the results and then take their union. Why do you not take the union
simultaneously with the parallel application of all rules (of a given layer), as
you are essentially doing for all applications of a single rule in Def. B.13? This
would get rid of the artificial sequential ordering, and the whole issue of con-
fluence would not even have to be raised as there is no non-determinism any
more.}

\reviewer{In the constructions defined in Appendix B, you ignore the required absence of
containment cycles. Might such cycles not appear in the union of the graphs,
even if the individual graphs are cycle-free?}

@@

\reviewer{I believe tr, T and R in this definition stand for the same thing.}
While the execution of the rules belonging to a layer happens in parallel, the execution of the layers of a transformation happens sequentially. As per \cref{def:transformation_step_semantics}, the input-output model $m_{inter}$ is the output of executing a given layer that is passed onto the next layer as input. Note that an empty input-output model ($\epsilon$) is passed as the second argument to the $layerstep$ relation in \cref{def:transformation_step_semantics}. This is because in \cref{def:layer_step_semantics} of layer step semantics, the second argument of the relation is used as an accumulator for the model fragments that are added to be added to the output of the previous layer once all the rules in the current layer have been executed. The transformation rules in a layer are expanded before execution ($l^{\star}$) such that polymorphism in the match elements can be handled (see \cref{def:transformation_rule_expansion_appendix}).

### 2016-06-29 text/old/appendix_abstraction_relation.tex
Move old sections. Divorce paper from definitions file.

\section{Validity and Completeness}
\label{subsec:abstraction_relation_validity_completeness}

In this section, we discuss the validity and completeness of the abstraction relation. Only proof sketches are presented here for \cref{prop:pc_validity} and \cref{lem:uniqueness}, while more complete proofs are found in \cref{sec:val_complete_path_cond_gen}, in the proofs of \cref{prop:pc_validity_appendix} and \cref{lem:uniqueness_appendix} respectively.

%\newpage
\begin{proposition}{(Validity)}
\label{prop:pc_validity}
Every path condition abstracts at least one transformation execution.
\end{proposition}
\begin{ps}

Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a DSLTans transformation. We wish to demonstrate that, for all path conditions $pc\in \textsc{Pathcond}(tr)$, there exists a transformation execution $ex\in \textsc{Exec}(tr)$ of the set of
rules used to build $pc$ such that $pc$ abstracts $ex$, as formally expressed in \cref{def:abstraction_pc_ex}. We can prove this property by induction on the set of transformations $\textsc{Transf}^{sr}_{tg}$ (see \cref{def:layer_transformation}), as follows:

\begin{itemize}
  \item \emph{Base case:} the base case is when $tr$ is the empty
  transformation. In this case, according to \cref{def:path_cond_gen} only the
  empty path condition $\epsilon_{pc}$ exists in the path condition set. We thus
  need to demonstrate that the empty path condition abstracts the empty
  transformation execution $\epsilon_{ex}$, as well as any execution for which
  the input model is never matched by any rule (consequently having an empty
  output model). For any of these transformation executions, Proposition~\ref{eq:abstr_input_output} of the abstraction relation definition is satisfied, as no rule copy exists in the path condition and the output of the transformation execution is empty -- empty typed graph homomorphisms thus satisfy all the conditions of the proposition. Proposition~\ref{eq:abstr_trace} of the abstraction relation definition also trivially holds because no traceability links exist either in the path condition or in any of the considered executions.

%   This is so because:
%   1) the empty typed graph homomorphism is an injective typed graph homomorphism
%   between the match part of the empty path condition and any input model; and 2) as
%   well as a surjective typed graph homomorphism between an empty output model
%   and the apply part of the empty path condition.

   \item \emph{Base case:} the base case is the case when we have $tr=[\;]$, i.e. the empty transformation. In this case, according to \cref{def:path_cond_gen_appendix}, only the empty path condition $\epsilon_{pc}$ exists in the path condition set. The empty path condition abstracts the empty transformation execution $\epsilon_{ex}$ (see \cref{def:modeltransformation_appendix}), as well as any execution for which the input model is never matched by any rule (consequently having an empty output model). For any of these transformation executions, \cref{eq:abstr_input_output} of the abstraction relation definition is satisfied, as: a) no rule copy exists in the path condition and the output of the transformation execution is empty -- empty typed graph homomorphisms thus satisfy the all the conditions of the proposition; and b) \cref{eq:abstr_trace} of the abstraction relation definition trivially holds because no traceability links exist either in the path condition or in any of the considered executions.

@@

In order to demonstrate the inductive case we need to show the property holds for all path conditions resulting from combining the rules of layer $l$ with any path condition generated for $tr$. These path conditions for transformation $tr'$ are built as expressed in \cref{def:path_cond_layer_comb}. According to this definition, path conditions for $tr'$ are built by incrementally combining the path conditions generated for $tr$ with a rule of layer $l$, until all the rules in $l$ have been treated. We can thus again use induction for this proof, this time on the set of possible layers $\textsc{Layer}^{sr}_{tg}$.

\begin{itemize}
  \item \emph{Base case:} this is the case where layer $l$ contains no rules. In this case, by the base case of \cref{def:path_cond_layer_comb}, no new path condition is added to the set of path conditions generated for the transformation $tr$. As such the $tr=tr'$ and by induction hypothesis the property trivially holds for all path conditions generated for $tr'$.

  \item \emph{Inductive case:} for the inductive case (transitive case of \cref{def:path_cond_layer_comb}) we need to show that, assuming the property holds for all path conditions generated for a transformation $tr$, then the property will also hold for a transformation $tr'$ -- where $tr'$ results from adding a new rule $rl$ to the last layer of $tr$. We will thus need to consider the four \reviewer{three?} cases of rule combination:\vspace{.2cm}

\begin{enumerate}
\item\label{lab:rule_case1} Rule $rl$ has no dependencies (\cref{def:rule_comb_no_dependencies}).
\item\label{lab:rule_case2} Rule $rl$ has dependencies and cannot execute (\cref{def:rule_comb_unsatisfied}).
\item\label{lab:rule_case3} Rule $rl$ has dependencies and may and/or will execute (\cref{def:rul_comb_partial_total}).
\end{enumerate}

The property trivially holds for case~\ref{lab:rule_case2}, given that no new path conditions are added to the path condition set generated for $tr$ and that the property holds for $tr$ by induction hypothesis. When a rule $rl$ is added to the last layer of $tr$ such that cases~\ref{lab:rule_case1} or~\ref{lab:rule_case3} occur, then the property can be shown to hold for $tr'$ as follows: 1) choose for a general path condition $pc$ generated for $tr$ an execution $ex$ such that $pc$ abstracts $ex$; 2) build an input model $m$ as the result of uniting the input model of $ex$ with a model that can be matched by $rl$; 3) execute $tr'$ having as input model $m$ to produce transformation execution $ex'$; and finally 4) demonstrate $ex'$ is abstracted by the path condition $pc'$ resulting from combining $pc$ with $rl$ whether rule $rl$ does not depend on $pc$ or rule $rl$ depends on $pc$ and may and/or will execute.

@@

\end{itemize}
\end{ps}

\begin{proposition}{(Completeness)}
\label{prop:pc_completeness}
Every transformation execution is abstracted by one path condition.
\end{proposition}
\begin{pf}
Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a DSLTans transformation. We wish to demonstrate that, for all transformation executions $ex\in \textsc{Exec}(tr)$, a path condition $pc\in \textsc{Pathcond}(tr)$ exists such that $ex$ is abstracted by $pc$, as formally expressed in \cref{def:abstraction_pc_ex}.\\

Completeness can be shown as a corollary of \cref{prop:pc_validity} about the \emph{validity} of path condition generation. The complete set of executions $\textsc{Exec}(tr)$ (see \cref{def:modeltransformation}) can be split into two kinds of executions:
\begin{enumerate}
  \item\label{lab:input_not_matched} The \emph{empty execution} $\epsilon_{ex}$ or the \emph{execution where the input model was not matched by any rule}. As mentioned in \cref{prop:pc_validity}, these executions are abstracted by the empty path condition $\epsilon_{pc}$.\vspace{.3cm}
  \item The \emph{execution $ex$ where a number of rules of $tr$ have been applied to the input model}, where each transformation rule $rl$ of $tr$ may have been applied more than once. In this case we have that, because all possible and valid rule combinations are considered when building path conditions, a path combination $pc$ exists that contains one or more copies of each of the rules used when operationally building $ex$. \\

  Moreover, during the proof of \emph{validity} of path condition generation in \cref{prop:pc_validity} we demonstrate that, when we add a new rule $rl$ to the last layer of a transformation $tr$ (such that we have a new transformation $tr'$), the rule combination step explained in Definitions~\ref{def:rule_comb_no_dependencies},~\ref{def:rule_comb_unsatisfied} and~\ref{def:rul_comb_partial_total} produces a new set of path conditions where each path condition in that set still abstracts at least one transformation execution of $tr'$. This part of the proof (the second induction) is achieved by building for transformation $tr'$ an input model $m$ that can be matched by $rl$ (as well as by all the other rules of $tr$), and then building from $m$ a new transformation execution that is abstracted by a path condition built for $tr'$. Because in the proof of \cref{prop:pc_validity}, $m$ is such that it can be matched by $rl$ an arbitrary amount of times, we know that, independently of the number of times a rule is applied during the construction of a transformation execution, a path condition abstracting that transformation execution exists.\\

Additionally, input elements that are not matched by any rule do not affect the abstraction relation, as explained in case~\ref{lab:input_not_matched} above. This means we also know that executions involving input models that are only partially matched by the rules of $tr$ are also abstracted by one path condition. 
\end{enumerate}

Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a DSLTans transformation. We wish to demonstrate that, for all transformation executions $ex\in \textsc{Exec}(tr)$, a path condition $pc\in \textsc{Pathcond}(tr)$ exists such that $ex$ is abstracted by $pc$, as formally expressed in \cref{def:abstraction_pc_ex_appendix}.
Completeness can be shown as a corollary of \cref{prop:pc_validity_appendix} about the \emph{validity} of path condition generation. The complete set of executions $\textsc{Exec}(tr)$ (see \cref{def:modeltransformation_appendix}) can be split into two kinds of executions:
\begin{enumerate}
  \item\label{lab:input_not_matched} The \emph{empty execution} $\epsilon_{ex}$ or the \emph{execution where the input model was not matched by any rule}. As mentioned in \cref{prop:pc_validity_appendix}, these executions are abstracted by the empty path condition $\epsilon_{pc}$.\vspace{.3cm}
  \item The \emph{execution $ex$ where a number of rules of $tr$ have been applied to the input model}, where each transformation rule $rl$ of $tr$ may have been applied more than once. In this case we have that, because all possible and valid rule combinations are considered when building path conditions, a path combination $pc$ exists that contains one or more copies of each of the rules used when operationally building $ex$. 
Moreover, during the proof of \emph{validity} of path condition generation in \cref{prop:pc_validity_appendix} we demonstrate that, when we add a new rule $rl$ to the last layer of a transformation $tr$ (such that we have a new transformation $tr'$), the rule combination step explained in Definitions~\ref{def:rule_comb_no_dependencies_appendix},~\ref{def:rule_comb_unsatisfied_appendix} and~\ref{def:rul_comb_partial_total_appendix} produces a new set of path conditions where each path condition in that set still abstracts at least one transformation execution of $tr'$. This part of the proof (the second induction) is achieved by building for transformation $tr'$ an input model $m$ that can be matched by $rl$ (as well as by all the other rules of $tr$), and then building from $m$ a new transformation execution that is abstracted by a path condition built for $tr'$. Because in the proof of \cref{prop:pc_validity_appendix}, $m$ is such that it can be matched by $rl$ an arbitrary amount of times, we know that, independently of the number of times a rule is applied during the construction of a transformation execution, a path condition abstracting that transformation execution exists.\\

Additionally, input elements that are not matched by any rule do not affect the abstraction relation, as explained in case~\ref{lab:input_not_matched} above. This means we also know that executions involving input models that are only partially matched by the rules of $tr$ are also abstracted by one path condition. 
\end{enumerate} 
\end{pf}

% Our path conditions construction algorithm takes all possible interactions of rules
% into account:
% \begin{enumerate}
% \item When building the path conditions for one layer we have built all
% the combinations of rules, as well as their possible interactions through
% \emph{disambiguation};
% \item When combining the path conditions from different layer different layers
% we have considered the cases where no interactions exist, or where backward
% links in rules requires certain certain elements to already exist in the
% generated model thus far.
% \end{enumerate}
% Note that, given the considered subset of DSLTrans constructs as described
% in \cref{sec:dsltrans}, no additional cases of rule interaction other than the
% ones described above need to be considered.
% From \cref{prop:pc_validity} we also know that at least one concrete
% transformation execution exists per path condition. In the proof of
% \cref{prop:pc_validity} we have built, for each path condition $A$,
% the simplest transformation execution that is abstracted by $A$. These
% transformation executions produce one instance of each element and association
% present in the apply pattern of $A$. However, given
% \cref{def:instance_pc_ex} in \cref{sec:proofs} (abstraction of
% a transformation execution by a path condition), an infinite amount of
% transformation executions are always abstracted by $A$. These executions
% correspond to the generation of output models holding an arbitrarily large
% amount of the elements and associations while still being abstracted by $A$.
% Because we have considered all possibilities of execution in our path condition
% construction algorithm, the union of all transformation executions built for
% each path condition constitutes the complete, infinite set of transformation
% executions. We thus know that every possible transformation execution is
% abstracted by at least one path condition.

% Note that we do not provide any guarantee that the same transformation execution
% cannot be abstracted by two different path conditions. Although given our symbolic
% execution assumptions we believe this to be the case, further mathematical
% exploration of this issue is necessary. Given however the fact that our
% verification algorithm exhaustively explores all path conditions, it is
% sufficient for the property proof to provide a \emph{no} result in the analysis
% of a path condition for the result of the verification of the property to be
% \emph{no}. \levi{keep this?}

\begin{lemma} (Uniqueness) 
\label{lem:uniqueness}
A transformation execution is abstracted by exactly one path condition.
\end{lemma}
\begin{ps}
Let $tr\in \textsc{Transf}^{sr}_{tr}$ be a model transformation. We will demonstrate that two different path conditions\\$pc_1, pc_2\in \textsc{Pathcond}(tr)$ cannot exist such that we have a transformation execution $ex\in \textsc{Exec}(tr)$ where $ex\Vvdash pc_1$ and $ex\Vvdash pc_2$.

We will do so by attempting to build an $ex\in \textsc{Exec}(tr)$ such that $ex\Vvdash pc_1$ and $ex\Vvdash pc_2$ and demonstrating that it is always the case that such is not possible. In order to structure our argumentation, we will consider two cases:
% \levi{the contradiction happens because at each point of the execution, when a rule is added any two path conditions from the existing set of path conditions will represent different executions. At all points if we consider two path conditions in the PC set we have that two different executions can be created if we find a model that satisfies the first condition of the abstraction relation, and it is always the case that none of those two executions can be abstracted by the two path conditions simultaneously.}
\begin{enumerate}
  \item\label{item:uniqueness_rule_no_dep} the case where no rules in $tr$ have dependencies.
  \item\label{item:uniqueness_rule_has_dep} the case where some rules in $tr$ have dependencies.
\end{enumerate}

We start by considering that $tr$ falls into case~\ref{item:uniqueness_rule_no_dep} above. By~\cref{def:path_cond_gen} of path condition generation, each rule appears at most once in a path condition. Also, by construction, each path condition always contains a different combination of rules. We additionally know from \cref{def:layer_transformation} that the rules that compose $tr$ necessarily have non-overlapping matchers. We can nonetheless build a model $m$ as the typed graph union of two input models $m_1$ and $m_2$, where injective typed graph morphisms can be found between the match parts of the rule copies that form $pc_1$, and $m_1$. Injective typed graph morphisms can be found as well between the match parts of the rules that form $pc_2$, and $m_2$. We thus know that injective typed graph morphisms can be found between the rule copies that compose $pc_1$ and $pc_2$, and $m$. This satisfies the first condition of \cref{eq:abstr_input_output} in ~\cref{def:abstraction_pc_ex} of abstraction relation.

@@

\end{ps}

>>>>

\section{Validity and Completeness of Path Condition Generation}
\label{sec:val_complete_path_cond_gen}
\counterwithin{proposition}{section}
\counterwithin{lemma}{section}
\counterwithin{equation}{section}
\setcounter{proposition}{0}
\setcounter{lemma}{0}
\setcounter{equation}{0}

\onehalfspacing

\begin{definition}{Path Condition\\}
\label{def:path_condition_appendix}
\CatchFileBetweenTags{\pcdef}{text/definitions}{pcdef}{\pcdef}
\end{definition}

\CatchFileBetweenTags{\pcappendix}{text/definitions}{pcappendix}{\pcappendix}

\begin{definition}{Combination of a Path Condition with a Rule}
\label{def:combine_pc_with_rule_appendix}
\CatchFileBetweenTags{\pccfrc}{text/definitions}{pccfrc}{\pccfrc}
\end{definition}

\CatchFileBetweenTags{\pccfrctextappendix}{text/definitions}{pccfrctextappendix}{\pccfrctextappendix}

\begin{definition}{Path Condition and Rule Combination -- No Dependencies\\}
\label{def:rule_comb_no_dependencies_appendix}
\CatchFileBetweenTags{\nodeps}{text/definitions}{nodeps}{\nodeps}
\end{definition}

\CatchFileBetweenTags{\nodepstextappendix}{text/definitions}{nodepstextappendix}{\nodepstextappendix}

\begin{definition}{Path Condition and Rule Combination -- Unsatisfied Dependencies\\} 
\label{def:rule_comb_unsatisfied_appendix}
\CatchFileBetweenTags{\unsatdeps}{text/definitions}{unsatdeps}{\unsatdeps}
\end{definition}

According to the pre-conditions of the equation presented in \cref{def:rule_comb_unsatisfied_appendix}, a path condition does not satisfy the dependencies present in a rule if there is no surjective typed graph homomorphism between the backward links of the rule and the symbolic traceability links of the path condition. Besides expressing the fact that all backward links must exist as symbolic traceability links the path condition, the surjective homomorphism allows modeling the case where dependencies expressed by two (or more) backward links between similarly typed elements can be satisfied one single symbolic traceability link in the path condition.

\begin{definition}{Single Partial and Total Combination of a Set of Path Conditions with a Rule\\}
\label{def:comb_path_cond_rule_single_appendix}
\CatchFileBetweenTags{\satdeps}{text/definitions}{satdeps}{\satdeps}

\begin{align}
\label{eq:pcomb_appendix}
\CatchFileBetweenTags{\satdepseqone}{text/definitions}{satdepseqone}{\satdepseqone}
\end{align}

\begin{align}
\label{eq:tcomb_appendix}
\CatchFileBetweenTags{\satdepseqtwo}{text/definitions}{satdepseqtwo}{\satdepseqtwo}
\end{align}

\end{definition}

Let us start by introducing relation $\stackrel{p\_comb}{\rightarrow}$, presented in \cref{eq:pcomb_appendix} of \cref{def:comb_path_cond_rule_single_appendix}. The relation takes as arguments a set of path conditions being accumulated for the current layer, the rule to be combined, and an $rl_{glue}$ argument indicating the place in each of the input path conditions the rule should be anchored to during the combination step. The relation's output is a new set of path conditions. This new set includes all the original path conditions, as well as each path condition in the accumulator set ``glued'' to a copy of rule being examined. Note that the relation $\stackrel{p\_total}{\rightarrow}$ in \cref{eq:tcomb_appendix} is similarly defined, except for the fact path conditions in the accumulator set are not preserved in the relation's output set.\\

\setcounter{equation}{0}

Let us now define how a rule is combined with a path condition, whenever its backward links can be found several times in that path condition. We formalize it in \cref{def:comb_path_cond_rule_mul_appendix}, by means of relations $\stackrel{p\_step}{\rightarrow}$ and $\stackrel{t\_step}{\rightarrow}$. These two relations operationally describe the sequence of steps necessary to ``glue'' a rule at multiples places of a path condition. The set of places targeted in the path condition for receiving a copy of the rule is given by the sets $partialSet$ and $totalSet$ (found respectively in \cref{eq:pstep} and \cref{eq:tstep}). As expected, these sets contain the set of traceability links in the path condition where copies of the rule need to be anchored to.

\begin{definition}{Multiple Partial and Total Combination of a Set of Path Conditions with a Rule\\}
\label{def:comb_path_cond_rule_mul_appendix}
\CatchFileBetweenTags{\satdepstwo}{text/definitions}{satdepstwo}{\satdepstwo}
\end{definition}

Having \cref{def:comb_path_cond_rule_single_appendix} and \cref{def:comb_path_cond_rule_mul_appendix} in mind, we can now proceed to define the complete combination relation of a rule with a path condition in the case of partially and totally satisfied dependencies.

\begin{definition}{Path Condition and Rule Combination -- Partially and Totally Satisfied Dependencies\\}
\label{def:rul_comb_partial_total_appendix}
\CatchFileBetweenTags{\satdepsthree}{text/definitions}{satdepsthree}{\satdepsthree}
\end{definition}

The top equation in \cref{def:rul_comb_partial_total_appendix} defines the $\stackrel{combine}{\rightarrow}$ relation for when rule $rl$ has dependencies that are satisfied by path condition $pc$. The pre-conditions in the equation state that the backward links in the rule are found in the path condition, as expected. Additionally, two sequential steps perform the gluing of the rule $rl$ on all path conditions in accumulator $AC$, wherever the rule is partially and/or totally found in each of those path conditions. Relations $\stackrel{p\_comb}{\rightarrow}$ and $\stackrel{t\_comb}{\rightarrow}$ presented in \cref{def:comb_path_cond_rule_mul_appendix} are used to model these two operational ``gluing'' steps. Functions $partialsat$ and $totalsat$, described in the latter part of \cref{def:rul_comb_partial_total_appendix}, are used to gather the places of path condition $pc$ where copies of the rule need to be anchored to.

\begin{definition} {Combining a Path Condition with a Layer\\}
\label{def:path_cond_layer_comb_appendix}
\CatchFileBetweenTags{\combpclayer}{text/definitions}{combpclayer}{\combpclayer}
\end{definition}

\CatchFileBetweenTags{\combpclayertextappendix}{text/definitions}{combpclayertextappendix}{\combpclayertextappendix}

\begin{definition} {Combining a Set of Path Conditions with a Layer\\}
\label{def:path_cond_set_layer_comb_appendix}
\CatchFileBetweenTags{\combpcsetlayer}{text/definitions}{combpcsetlayer}{\combpcsetlayer}
\end{definition}

\CatchFileBetweenTags{\combpcsetlayertextappendix}{text/definitions}{combpcsetlayertextappendix}{\combpcsetlayertextappendix}

\begin{definition} {Path Condition Generation\\}
\label{def:path_cond_gen_appendix}
\CatchFileBetweenTags{\pathcondgen}{text/definitions}{pathcondgen}{\pathcondgen}
\end{definition}

\CatchFileBetweenTags{\pathcondgentextappendix}{text/definitions}{pathcondgentextappendix}{\pathcondgentextappendix}

\begin{definition}{Abstraction of a Transformation Execution by a Path Condition\\}
\label{def:abstraction_pc_ex_appendix}
\CatchFileBetweenTags{\abstractionrelappendix}{text/definitions}{abstractionrelappendix}{\abstractionrelappendix}
\end{definition}

To understand the abstraction relation in \cref{def:abstraction_pc_ex_appendix}, recall
that during the construction of a transformation execution rules are matched
injectively in the input model. This information is present in the first
condition of the abstraction relation (Proposition~\ref{eq:abstr_input_output_appendix})
via the injective typed graph homomorphism between the match part of the copies
of rules ``glued'' onto the path condition and the containment transitive
closure of the input part of the transformation execution. This relation
enforces the fact that certain parts of the execution were found, or matched, by
certain parts of the path condition.
On the other hand, the surjection from the output of the execution towards the
apply part of the path condition models the fact that the output of the
execution has been completely built by instantiating the apply parts of the
rules contained in the path condition.

The second condition of the abstraction relation
(Proposition~\ref{eq:abstr_trace_appendix}) checks for the fact that symbolic
traceability links in the path condition and traceability links in the execution
correctly correspond to each other. This is modeled by the fact that all
strongly connected components in the path condition, composed only of symbolic
traceability links, are injectively found on the execution. This injection
models the fact that traceability graphs between individual or combined rules in
the path condition are necessarily found in the execution. Note that components
of the path condition are considered because of the fact that disconnected rules
in the path condition may have matched over common elements of a particular
execution. As such, a full injection between the complete traceability graph in
the path condition and the execution would be incorrect. Additionally, in the
second part of Proposition~\ref{eq:abstr_trace_appendix} we check the fact that every
traceability link in the execution can be found in the path condition. This
additional sanity check enforces that no spurious traceability links that could
not have been created by the rules present in the path condition exist in the
transformation execution.

Finally, the last clause of the abstraction relation states that rule copies
that are repeated a number of times in the path condition need to be found at
least a similar amount of times in the abstracted transformation execution.

It is important to mention that another abstraction relation, weaker or
stronger, could have been chosen. The abstraction relation presented in
\cref{def:abstraction_pc_ex_appendix} suits our needs in the sense that it allows us to
demonstrate the validity and completeness of our proof technique, as we will
show in the text follows. Additionally, it is particularly interesting because
it makes sure that, given a DSLTrans transformation, each of its transformation
executions is abstracted by one and only one of its path conditions. This result
adds to the consistency of our theory and is also exposed later in this section.

\begin{proposition}{(Validity)}
\label{prop:pc_validity_appendix}
Every path condition abstracts at least one transformation execution.
\end{proposition}
\begin{pf}
Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a DSLTrans transformation. We wish to demonstrate that, for all path conditions $pc\in \textsc{Pathcond}(tr)$, there exists a transformation execution $ex\in \textsc{Exec}(tr)$ of the set of rules used to build $pc$ such that $pc$ abstracts $ex$ (i.e. $ex\Vvdash pc$), as formally expressed in \cref{def:abstraction_pc_ex_appendix}. We can prove this property by induction on the set of transformations $\textsc{Transf}^{sr}_{tg}$ (see \cref{def:layer_transformation_appendix}), as follows:

% We will start by showing that the property
% holds for the path conditions that are generated for one single layer of a
% DSLTrans transformation, which in what follows we will call lemma 1.
% In this case the inductive proof will require the following steps:

\begin{itemize}
  \item \emph{Base case:} the base case is the case when we have $tr=[\;]$, i.e. the empty transformation. In this case, according to \cref{def:path_cond_gen_appendix}, only the empty path condition $\epsilon_{pc}$ exists in the path condition set. The empty path condition abstracts the empty transformation execution $\epsilon_{ex}$ (see \cref{def:modeltransformation_appendix}), as well as any execution for which the input model is never matched by any rule (consequently having an empty output model). For any of these transformation executions, \cref{eq:abstr_input_output} of the abstraction relation definition is satisfied, as: a) no rule copy exists in the path condition and the output of the transformation execution is empty -- empty typed graph homomorphisms thus satisfy the all the conditions of the proposition; and b) \cref{eq:abstr_trace} of the abstraction relation definition trivially holds because no traceability links exist either in the path condition or in any of the considered executions.

@@

In order to demonstrate the inductive case we need to show the property holds for all path conditions resulting from combining the rules of layer $l$ with any path condition generated for $tr$. These path conditions for transformation $tr'$ are built as expressed in \cref{def:path_cond_layer_comb_appendix}. According to this definition, path conditions for $tr'$ are built by incrementally combining the path conditions generated for $tr$ with a rule of layer $l$, until all the rules in $l$ have been treated. We can thus again use induction for this proof, this time on the set of possible layers $\textsc{Layer}^{sr}_{tg}$ built as expressed in \cref{def:layer_transformation_appendix}.

\begin{itemize}
  \item \emph{Base case:} this is the case where layer $l$ contains no rules. In this case, by the base case of definition~\ref{def:path_cond_layer_comb_appendix}, no new path condition is added to the set of path conditions generated for the transformation $tr$. As such the $tr=tr'$ and by induction hypothesis the property trivially holds for all path conditions generated for $tr'$.

  \item \emph{Inductive case:} for the inductive case (transitive case of \cref{def:path_cond_layer_comb_appendix}) we need to show that, assuming the property holds for all path conditions generated for a transformation $tr$, then the property will also hold for a transformation $tr'$ -- where $tr'$ results from adding a new rule $rl$ to the last layer of $tr$. We will thus need to consider the four cases of rule combination:\vspace{.2cm}

\begin{enumerate}
\item\label{lab:rule_case1_appendix} Rule $rl$ has no dependencies (\cref{def:rule_comb_no_dependencies_appendix}).
\item\label{lab:rule_case2_appendix} Rule $rl$ has dependencies and cannot execute (\cref{def:rule_comb_unsatisfied_appendix}).
\item\label{lab:rule_case3_appendix} Rule $rl$ has dependencies and may and/or will execute (\cref{def:rul_comb_partial_total_appendix}).
\end{enumerate}

@@

\end{itemize}
\end{pf}

\begin{proposition}{(Completeness)}
\label{prop:pc_completeness_appendix}
Every transformation execution is abstracted by one path condition.
\end{proposition}
\CatchFileBetweenTags{\pathgencompletenessproofappendix}{text/definitions}{pathgencompletenessproofappendix}{\pathgencompletenessproofappendix}

\begin{lemma} (Uniqueness)
\label{lem:uniqueness_appendix}
A transformation execution is abstracted by exactly one path condition.
\end{lemma}
\begin{pf}
Let $tr\in \textsc{Transf}^{sr}_{tr}$ be a model transformation. We will demonstrate that two different path conditions $pc_1, pc_2\in \textsc{Pathcond}(tr)$ cannot exist such that we have a transformation execution $ex\in \textsc{Exec}(tr)$ where $ex\Vvdash pc_1$ and $ex\Vvdash pc_2$.

We will do so by attempting to to build an $ex\in \textsc{Exec}(tr)$ such that $ex\Vvdash pc_1$ and $ex\Vvdash pc_2$ and demonstrating that it is always the case that such is not possible. In order to structure our argumentation, we will consider two cases:
% \levi{the contradiction happens because at each point of the execution, when a rule is added any two path conditions from the existing set of path conditions will represent different executions. At all points if we consider two path conditions in the PC set we have that two different executions can be created if we find a model that satisfies the first condition of the abstraction relation, and it is always the case that none of those two executions can be abstracted by the two path conditions simultaneously.}
\begin{enumerate}
  \item\label{item:uniqueness_rule_no_dep_appendix} the case where no rules in $tr$ have dependencies.
  \item\label{item:uniqueness_rule_has_dep_appendix} the case where some rules in $tr$ have dependencies.
\end{enumerate}

We start by considering that $tr$ falls into case~\ref{item:uniqueness_rule_no_dep_appendix} above. By \cref{def:path_cond_gen_appendix} of path condition generation, each rule appears at most once in a path condition. Also, by construction, each path condition always contains a different combination of rules. We additionally know from \cref{def:layer_transformation_appendix} that the rules that compose $tr$ necessarily have non-overlapping matchers. We can nonetheless build a model $m$ as the typed graph union of two input models $m_1$ and $m_2$, where injective typed graph morphisms can be found between the match parts of the rule copies that form $pc_1$, and $m_1$. Injective typed graph morphisms can be found as well between the match parts of the rules that form $pc_1$, and $m_2$. We thus know that injective typed graph morphisms can be found between the rule copies that compose $pc_1$ and $pc_2$, and $m$. This satisfies the first condition of \cref{eq:abstr_input_output} in ~\cref{def:abstraction_pc_ex_appendix} of abstraction relation.

@@

\end{pf}

\clearpage
